```{r knitr_setup, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Data types

_1) Provide a URL to the dataset._

I downloaded my dataset from https://query.data.world/s/7iysjzfn22qo5cgarzul2tv3wh6f3a
https://www.kaggle.com/fmejia21/demographics-of-academy-awards-oscars-winners/data#Oscars-demographics-DFE.csv

_2) Explain why you chose this dataset._ 

I am interested in seeing how much the diversity of academy award winners has changed over the years, if it did.

_3) What are the entities in this dataset? How many are there?_
c
Entities are Oscar winners' names. There are 441.

_4) How many attributes are there in this dataset?_

There are 27 attributes.

_5) What is the datatype of each attribute (categorical -ordered or unordered-, numeric -discrete or continuous-, datetime, geolocation, other)? Write a short sentence stating how you determined the type of each attribute. Do this for at least 5 attributes, if your dataset contains more than 10 attributes, choose 10 of them to describe._

| Num | Name | Type | Description |
|-----|------|------|-------------|
| 1 | `birthplace` | categorical unordered | Identifier of birthplace which takes values from finite set of possible birthplaces |
| 2 | `date_of_birth` | DateTime | Birth dates are string values measured in time units |
| 3 | `race_ethnicity` | categorical unordered | Can take value from finite set of possible races |
| 4 | `religion` | categorical unordered | Can take value from finite set of possible religions |
| 5 | `sexual_orientation` | categorical unordered | Identifies sexual orientation, taking values from a finite set of possible sexual orientations |
| 6 | `year_of_award` | numerical discrete | Specifies exact year of receipt of award, which doesn't change |
| 7 | `award` | categorical unordered | Street address of arrest |
| 8 | `movie` | categorical unordered | Can take value from finite set of movies |
| 9 | `person` | categorical unordered | Identifies person's name from a finite set |
| 10 | `biourl` | categorical unordered | Can take value from finite set of URLs |

_6) Write R code that loads the dataset using function `read_csv`. Were you able to load the data successfully? If no, why not?_
For this site, I couldn't find a .csv URL so as a result I had to download the .csv file. The only way I could execute a command that properly pulls up the dataset is if I called read.csv and passed in the path where the downloaded file is as an argument (in this case my downloads folder). 
```{r load_data, message=FALSE}
library(tidyverse)

Oscars_Demographics <- read.csv("~/Downloads/Oscars-demographics-DFE.csv")

Oscars_Demographics %>% slice(1:10)
```

## Wrangling

1) My pipeline finds and groups the earliest and most recent years of Oscar winners based on race, and orders the data based on which racial group won an award first.
```{r pipeline}
earliest_award <- Oscars_Demographics %>%
  select(person, race_ethnicity, sexual_orientation, award, year_of_award) %>%
  group_by(race_ethnicity) %>%
  summarize(earliest=min(year_of_award), latest=max(year_of_award)) %>%
  arrange(earliest)
earliest_award
```

## Plotting

1) This barplot shows the average arrest age per district (ignoring ages <= 0)

```{r plot}
mean_ages %>%
  ggplot(aes(x=district, y=mean_age)) +
    geom_bar(stat="identity") +
    coord_flip()
```