---
title: 'Project 2: Wrangling and Exploratory Data Analysis'
author: "Sonya Radichkova"
date: "4/1/2020"
output: html_document
---
Connecting to the database and loading r libraries
```{r connect_db, include=FALSE}
db <- DBI::dbConnect(RSQLite::SQLite(), "~/Downloads/lahman2016.sqlite")
library(rvest)
library(tidyverse)
library(lubridate)
```

## Wrangling
Problem 1:
To compute total payroll and winning percentage, we must group by teamID and yearID in order to
ensure payroll is grouped correctly for its corresponding team and year. I joined the Salaries and Teams tables on team and yearid so the final table is grouped correctly and so I could access the Team table's wins and number of games to calculate into the winning percentage.

```{sql run_query, echo=TRUE, connection=db, eval=TRUE, output.var="payroll_df"}
SELECT s.teamID, s.yearID, SUM(s.salary) as total_payroll, t.W, t.G, (cast(t.W as float) / cast(t.G as float)) * 100 as winning_percent, s.salary as sal, AVG(s.salary) as mean_payroll
FROM Salaries as s
INNER JOIN Teams as t
ON t.teamID = s.teamID AND t.yearID = s.yearID
WHERE s.yearID >=1990 AND s.yearID<=2014
GROUP BY s.teamID , s.yearID, t.W, t.G
ORDER BY s.teamID, s.yearID
```



## Exploratory Data Analysis
# Payroll distribution
Problem 2:
```{r}
payroll_df %>%
  ggplot(mapping=aes(x=yearID, y=total_payroll, group=yearID)) +
  geom_boxplot() +
  labs(title = "Payroll Distribution") + ylab("Total Payroll") + xlab("Year")
  
```


Question 1:
Over time, the boxplot shows a more of a spread between the minimum and maximum along with the quartiles of each box becoming more spread out over time. There appear to be some more outliers starting from the 2000s that go beyond the box and whisker plots. The distribution of payrolls in the 2000s is more spread out than it was in the 1990s, so the spread will likely increase beyond 2015, while the central tendency of the data will increase at a very slow rate.

Problem 3:
I used the standard deviation function in order to observe the spread of teams' payrolls, which shows that there is still a relative increase in the spread of payroll among team players.
```{r}
payroll_df %>%
  group_by(yearID) %>%
  summarize(mean_pay=sd(sal)) %>%
  ggplot(mapping=aes(x=yearID, y=mean_pay)) +
  geom_bar(stat="identity") +
  labs(title = "Payroll Spread") + ylab("Standard Deviation") + xlab("Year")
```


# Correlation Between Payroll and Winning Percentage
Problem 4:
I used mutate to create a new column with the year breaks, then made sure the data was grouped together by team and that discrete year group, and summarized average winning percent and average payroll based on those groupings in order to plot the data shown below.
```{r}
highlight <- payroll_df %>% 
  filter(teamID=="OAK")
  payroll_df %>%
  mutate(discrete_year=cut(yearID, breaks=5)) %>%
  group_by(teamID, discrete_year) %>%
  summarize(avg_winning_percent=mean(winning_percent), average_payroll=mean(mean_payroll)) %>%
  ggplot(mapping=aes(x=average_payroll, y=avg_winning_percent)) +
  facet_grid(cols=vars(discrete_year)) +
  geom_point() +

  geom_smooth(method=lm)
  
```

Question 2:
Team payrolls have significantly spread out over time, and only later on does it show that paying more does show a slight trend of higher winning percentage, however, Atlanta Braves had the highest percent with NYA, however unlike NYA, the Atlanta Braves didn't necessarily pay as much. Looking at the Oakland A's spending efficiency, their highest payroll average was right after the moneyball period, after they hit their peak winning percent, so as the payroll average went up after moneyball, their winning percentage decreased, which isn't as efficient in terms of pay. And once the payroll average went down a bit between 2009-2014, their winning percentage slightly went up again. This goes to show that there's not exactly a direct correlation between a higher payroll and a higher winning percentage. 

## Data Transformations
# Standardization across years
Problem 5:
I directly modified the payroll dataframe to include a new column of the new formula which is the salary minus average payroll, and the data frame is already grouped on team and year so the calculations should be based on the given team and year, all of this divided by the standard deviation of the year as per the standardization formula.
```{r}
payroll_df<-payroll_df %>%
  mutate(standardized_payroll=(sal-mean_payroll)/sd(year)) 
head(payroll_df)
```

Problem 6
I specifically included breaks and labels in the mutate function because the grouped years weren't showing up the same as number 4 so I maually chose where to do the breaks and labeled them the way they were before to fix that issue. I made the stand_pay variable the mean of the standardized payroll (since a team will likely have more than 1 piece of data within a 5 year time period) in order for the standardized payroll to fall within its respective time frame.


```{r}
payroll_df %>%
  mutate(discrete_year=cut(yearID, breaks=c(1989, 1995, 2000, 2004, 2009, 2014), labels = c("(1990,1995]","(1995,2000]","(2000,2004]", "(2004,2009]", "(2009,2014]")))  %>%
  group_by(teamID, discrete_year) %>%
  summarize(avg_winning_percent=mean(winning_percent), stand_pay=mean(standardized_payroll)) %>%
  ggplot(mapping=aes(x=stand_pay, y=avg_winning_percent)) +
  facet_grid(cols=vars(discrete_year)) +
  geom_point() +

  geom_smooth(method=lm)
```
 Question 3:
 Seeing number 6's plot compared to number 4's shows the trendline going in the complete opposite direction, going downwards rather than upwards, and that the average winning percent doesn't appear to directly correlate to standardized payroll at all. There is less of a spread in the data, but the majority of the plots lie near the direct center, with the exception of the moneyball period, where winning percent was much more spread out. Through the course of these time periods, pay appears to increase while the winning percentage appears to slowly decrease.

#Expected Wins
Problem 7

#Spending Efficiency


```{r disconnectdb, include=FALSE}
DBI::dbDisconnect(db)
```
